DateLife: leveraging databases and analytical tools to reveal the dated Tree of Life

### 2022-07-06 dryad/README.Rmd
fixing indentation; renaming jpg files

	Corresponding Researcher name: Luna L. Sanchez Reyes
  Corresponding Researcher institution: University of California, Merced, USA
  Corresponding Researcher email: sanchez.reyes.luna@gmail.com

	Co-researcher 1 name: Emily Jane McTavish
	Co-researcher 1	institution: University of California, Merced, USA

	Co-researcher 2 name: Brian C. O'Meara
	Co-researcher 2	institution: University of Tennessee, Knoxville, USA

@@

```
cp tables/table-fringillidae-all.pdf dryad/Supplementary_Table_S1.pdf
cp tables/table-fringillidae-all-summary.pdf dryad/Supplementary_Table_S2.pdf
cp figures/figure-cross-validation/fig-cross-validation-xy-plots-diffs.pdf dryad/Supplementary_Figure_S1.pdf
cp figures/figure-cross-validation/cross_validation_1.jpg dryad/Supplementary_Figure_S2.jpg
cp figures/figure-cross-validation/cross_validation_2.jpg dryad/Supplementary_Figure_S3.jpg
cp figures/figure-cross-validation/cross_validation_3.jpg dryad/Supplementary_Figure_S4.jpg
cp figures/figure-cross-validation/cross_validation_4.jpg dryad/Supplementary_Figure_S5.jpg
cp figures/figure-cross-validation/cross_validation_5.jpg dryad/Supplementary_Figure_S6.jpg
cp figures/figure-cross-validation/cross_validation_6.jpg dryad/Supplementary_Figure_S7.jpg
cp figures/figure-cross-validation/cross_validation_7.jpg dryad/Supplementary_Figure_S8.jpg
cp figures/figure-cross-validation/cross_validation_8.jpg dryad/Supplementary_Figure_S9.jpg
cp figures/figure-cross-validation/cross_validation_9.jpg dryad/Supplementary_Figure_S10.jpg
cp figures/figure-cross-validation/cross_validation_10.jpg dryad/Supplementary_Figure_S11.jpg
cp figures/figure-cross-validation/cross_validation_11.jpg dryad/Supplementary_Figure_S12.jpg
cp figures/figure-cross-validation/cross_validation_12.jpg dryad/Supplementary_Figure_S13.jpg
cp figures/figure-cross-validation/cross_validation_13.jpg dryad/Supplementary_Figure_S14.jpg
cp figures/figure-cross-validation/cross_validation_14.jpg dryad/Supplementary_Figure_S15.jpg
cp figures/figure-cross-validation/cross_validation_15.jpg dryad/Supplementary_Figure_S16.jpg
cp figures/figure-cross-validation/cross_validation_16.jpg dryad/Supplementary_Figure_S17.jpg
cp figures/figure-cross-validation/cross_validation_17.jpg dryad/Supplementary_Figure_S18.jpg
cp figures/figure-cross-validation/cross_validation_18.jpg dryad/Supplementary_Figure_S19.jpg
cp figures/figure-cross-validation/cross_validation_19.jpg dryad/Supplementary_Figure_S20.jpg
```

>>>>

	Corresponding Researcher   
	name: Luna L. Sanchez Reyes    
  institution: University of California, Merced, USA    
  email: sanchez.reyes.luna@gmail.com

	Co-researcher 1    
	name: Emily Jane McTavish    
	institution: University of California, Merced, USA

	Co-researcher 2    
	name: Brian C. O'Meara    
	institution: University of Tennessee, Knoxville, USA

@@

```
cp tables/table-fringillidae-all.pdf dryad/Supplementary_Table_S1.pdf
cp tables/table-fringillidae-all-summary.pdf dryad/Supplementary_Table_S2.pdf
cp figures/figure-cross-validation/fig-cross-validation-xy-plots-diffs.pdf dryad/Supplementary_Figure_S1.pdf
```

### 2022-07-06 manuscript/manuscript.Rmd
fixing indentation; renaming jpg files

abstract: |
  Achieving a high-quality reconstruction of a phylogenetic tree with branch lengths proportional to absolute time (chronogram) is a difficult and time-consuming task. But the increased availability of fossil and molecular data, and time-efficient analytical techniques has resulted in many recent publications of large chronograms for a large number and wide diversity of organisms.
    Knowledge of the evolutionary time frame of organisms is key for research in the natural sciences. It also represent valuable information for education, science communication, and policy decisions.
  When chronograms are shared in public, open databases this wealth of expertly-curated and peer-reviewed data on evolutionary timeframe is exposed in a programatic and reusable way, as intensive and localized efforts have improved data sharing practices, as well as incentivizited open science in biology.
  Here we present DateLife, a service implemented as an R package and an R Shiny website application available at www.datelife.org, that provides functionalities for efficient and easy finding, summary, reuse, and reanalysis of expert, peer-reviewed, public data on time frame of evolution.
  The main DateLife workflow constructs a chronogram for any given combination of taxon names by searching a local chronogram database constructed and curated from the Open Tree of Life Phylesystem phylogenetic database, which incorporates phylogenetic data from the TreeBASE database as well.
  We implement and test methods for summarizing time data from multiple source chronograms using supertree and congruification algorithms, and using age data extracted from source chronograms as secondary calibration points to add branch lengths proportional to absolute time to a tree topology.
  DateLife will be useful to increase awereness of the existing variation in alternative hypothesis of evolutionary time for the same organisms, and can foster exploration of the effect of alternative evolutionary timing hypotheses on the results of downstream analyses, providing a framework for a more informed interpretation of evolutionary results.

>>>>

abstract: |
  Achieving a high-quality reconstruction of a phylogenetic tree with branch lengths proportional to absolute time (chronogram) is a difficult and time-consuming task. But the increased availability of fossil and molecular data, and time-efficient analytical techniques has resulted in many recent publications of large chronograms for a large number and wide diversity of organisms.
    Knowledge of the evolutionary time frame of organisms is key for research in the natural sciences. It also represent valuable information for education, science communication, and policy decisions.
  When chronograms are shared in public and open databases, this wealth of expertly-curated and peer-reviewed data on evolutionary timeframe is exposed in a programatic and reusable way, as intensive and localized efforts have improved data sharing practices, as well as incentivizited open science in biology.
  Here we present DateLife, a service implemented as an R package and an R Shiny website application available at www.datelife.org, that provides functionalities for efficient and easy finding, summary, reuse, and reanalysis of expert, peer-reviewed, public data on time frame of evolution.
  The main DateLife workflow constructs a chronogram for any given combination of taxon names by searching a local chronogram database constructed and curated from the Open Tree of Life Phylesystem phylogenetic database, which incorporates phylogenetic data from the TreeBASE database as well.
  We implement and test methods for summarizing time data from multiple source chronograms using supertree and congruification algorithms, and using age data extracted from source chronograms as secondary calibration points to add branch lengths proportional to absolute time to a tree topology.
  DateLife will be useful to increase awareness of the existing variation in alternative hypothesis of evolutionary time for the same organisms, and can foster exploration of the effect of alternative evolutionary timing hypotheses on the results of downstream analyses, providing a framework for a more informed interpretation of evolutionary results.

### 2022-07-06 manuscript/manuscript.tex
fixing indentation; renaming jpg files

\abstract{%
Achieving a high-quality reconstruction of a phylogenetic tree with branch lengths proportional to absolute time (chronogram) is a difficult and time-consuming task. But the increased availability of fossil and molecular data, and time-efficient analytical techniques has resulted in many recent publications of large chronograms for a large number and wide diversity of organisms.
Knowledge of the evolutionary time frame of organisms is key for research in the natural sciences. It also represent valuable information for education, science communication, and policy decisions.
When chronograms are shared in public, open databases this wealth of expertly-curated and peer-reviewed data on evolutionary timeframe is exposed in a programatic and reusable way, as intensive and localized efforts have improved data sharing practices, as well as incentivizited open science in biology.
Here we present DateLife, a service implemented as an R package and an R Shiny website application available at www.datelife.org, that provides functionalities for efficient and easy finding, summary, reuse, and reanalysis of expert, peer-reviewed, public data on time frame of evolution.
The main DateLife workflow constructs a chronogram for any given combination of taxon names by searching a local chronogram database constructed and curated from the Open Tree of Life Phylesystem phylogenetic database, which incorporates phylogenetic data from the TreeBASE database as well.
We implement and test methods for summarizing time data from multiple source chronograms using supertree and congruification algorithms, and using age data extracted from source chronograms as secondary calibration points to add branch lengths proportional to absolute time to a tree topology.
DateLife will be useful to increase awereness of the existing variation in alternative hypothesis of evolutionary time for the same organisms, and can foster exploration of the effect of alternative evolutionary timing hypotheses on the results of downstream analyses, providing a framework for a more informed interpretation of evolutionary results.
}

>>>>

\abstract{%
Achieving a high-quality reconstruction of a phylogenetic tree with branch lengths proportional to absolute time (chronogram) is a difficult and time-consuming task. But the increased availability of fossil and molecular data, and time-efficient analytical techniques has resulted in many recent publications of large chronograms for a large number and wide diversity of organisms.
Knowledge of the evolutionary time frame of organisms is key for research in the natural sciences. It also represent valuable information for education, science communication, and policy decisions.
When chronograms are shared in public and open databases, this wealth of expertly-curated and peer-reviewed data on evolutionary timeframe is exposed in a programatic and reusable way, as intensive and localized efforts have improved data sharing practices, as well as incentivizited open science in biology.
Here we present DateLife, a service implemented as an R package and an R Shiny website application available at www.datelife.org, that provides functionalities for efficient and easy finding, summary, reuse, and reanalysis of expert, peer-reviewed, public data on time frame of evolution.
The main DateLife workflow constructs a chronogram for any given combination of taxon names by searching a local chronogram database constructed and curated from the Open Tree of Life Phylesystem phylogenetic database, which incorporates phylogenetic data from the TreeBASE database as well.
We implement and test methods for summarizing time data from multiple source chronograms using supertree and congruification algorithms, and using age data extracted from source chronograms as secondary calibration points to add branch lengths proportional to absolute time to a tree topology.
DateLife will be useful to increase awareness of the existing variation in alternative hypothesis of evolutionary time for the same organisms, and can foster exploration of the effect of alternative evolutionary timing hypotheses on the results of downstream analyses, providing a framework for a more informed interpretation of evolutionary results.
}
